---
title: "Canvas Summarize and Explore Grades Across Classes"
author: "Nathan Craig"
date: "13 May 2021"
output:
  html_document:
    toc: yes
    number_sections: yes
    df_print: paged
    link-citations: yes
bibliography: [references.bib, packages.bib]
---

# Introduction
This document illustrates some Exploratory Data Analysis (EDA) [@tukey1977] of Canvas exam grades. The data are wrangled with `tidyverse` and `dplyr` [@R-tidyverse] functions, then plotted with `ggplot2` [@R-ggplot2] and `ggstatsplot` [@R-ggstatsplot].

# Load Libraries
We begin by loading the necessary libraries.
```{r message=FALSE, warning=FALSE}

# Read CSV file
library(readr)

# Data wrangling and formatting
library(tidyverse)
library(kableExtra)
library(knitr)
opts_chunk$set(comment=NA,
               prompt=FALSE,
               cache=FALSE,
               echo=TRUE,
               results='asis')

# Stats
library(summarytools)

# Plotting
library(ggplot2)
library(ggstatsplot)
```

# Load Data and Wrangle
Since we are working with multiple classes, in this case eight comparable instances, the files were[ pulled from canvas](https://n8craig.github.io/nmsu-grades/Canvas_Grades_Get_Grades_Multiple_Classes.html) and then saved to a csv file. Here that file is read from disk and wrangled into shape.

```{r}
gb_list <- read_csv("data/gb_list.csv")

gb_list <- gb_list %>% 
  mutate(course_id = as_factor(course_id)) %>% 
  mutate(course_name = as_factor(course_name))
```

The base R `factor()` function is applied to order the classes chronologically. This ensures proper ordering in the graphs that follow.
```{r}
# order the factors properly
gb_list$course_name <- factor(gb_list$course_name,
                              levels= c(
                                "ANTH-125G S19",
                                "ANTH-125G F19",
                                "HON-235G F19",
                                "ANTH-125G S20",
                                "ANTH-125G F20",
                                "HON-235G F20",
                                "ANTH-1140G S21",
                                "HNRS-2161G S21"
                              ))
```

# Exam 1

```{r}
library(kableExtra)
```



```{r}
gb_list %>%
  filter(str_detect(assignment_name, "^Exam1|^EXAM1")) %>%
  group_by(course_name) %>%
  select(assignment_name, score, course_name) %>%
  descr(stats = c("mean", "sd", "min", "med", "max", "skewness", "kurtosis"),
        transpose = TRUE,
        headings = FALSE
        ) %>% 
  tb() %>% 
  kable(format = "html", digits = 2) %>% 
  kable_styling(bootstrap_options = "striped")
```


```{r}
gb_list %>% 
  filter(str_detect(assignment_name, "^Exam1|^EXAM1")) %>% 
  group_by(course_name) %>% 
ggplot(., aes(x = score))+
  geom_histogram()+
  facet_wrap(~ course_name)+
    labs(
    title = "Exam 1 Histograms Across Classes"
  )
```

```{r}
gb_list %>% 
  group_by(course_id) %>% 
  filter(str_detect(assignment_name, "^Exam1|^EXAM1")) %>% 
  ggplot(.,aes(x=course_name, y = score, group=course_name))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    labs(
    title = "Exam 1 Scores Across Different Classes",
    x = "Course",
    y = "Exam Score"
  )
```

```{r}
gb_list %>% 
  group_by(course_name) %>% 
  filter(str_detect(assignment_name, "^Exam1|^EXAM1")) %>%
  ggbetweenstats(
    data = .,
    x = course_name,
    y = score,
    pairwise.comparisons = FALSE)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    labs(
    title = "Exam 1 Scores Across Different Classes",
    x = "Course",
    y = "Exam Score"
  )
```

# Exam 2

```{r}
gb_list %>%
  filter(str_detect(assignment_name, "^Exam2|^EXAM2")) %>%
  group_by(course_name) %>%
  select(assignment_name, score, course_name) %>%
  descr(stats = c("mean", "sd", "min", "med", "max", "skewness", "kurtosis"),
        transpose = TRUE,
        headings = FALSE) %>% 
  tb() %>% 
  kable(format = "html", digits = 2) %>% 
  kable_styling(bootstrap_options = "striped")
```


```{r}
gb_list %>% 
  filter(str_detect(assignment_name, "^Exam2|^EXAM2")) %>% 
  group_by(course_name) %>% 
ggplot(., aes(x = score))+
  geom_histogram()+
  facet_wrap(~ course_name)+
  labs(
    title = "Exam 2 Histograms Across Classes"
  )
```

```{r}
gb_list %>% 
  group_by(course_id) %>% 
  filter(str_detect(assignment_name, "^Exam2|^EXAM2")) %>% 
  ggplot(.,aes(x=course_name, y = score, group=course_name))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(
    title = "Exam 2 Scores Across Different Classes",
    x = "Course",
    y = "Exam Score"
  )
```

```{r}
gb_list %>% 
  group_by(course_name) %>% 
  filter(str_detect(assignment_name, "^Exam2|^EXAM2")) %>%
  ggbetweenstats(
    data = .,
    x = course_id,
    y = score,
    pairwise.comparisons = FALSE)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    labs(
    title = "Exam 2 Scores Across Different Classes",
    x = "Course",
    y = "Exam Score"
  )
```

# Exam 3

```{r}
gb_list %>%
  filter(str_detect(assignment_name, "^Exam3|^EXAM3")) %>%
  group_by(course_name) %>%
  select(assignment_name, score, course_name) %>%
  descr(stats = c("mean", "sd", "min", "med", "max", "skewness", "kurtosis"),
        transpose = TRUE,
        headings = FALSE) %>% 
  tb() %>% 
  kable(format = "html", digits = 2) %>% 
  kable_styling(bootstrap_options = "striped")
```

```{r}
gb_list %>% 
  filter(str_detect(assignment_name, "^Exam3|^EXAM3|^Fin|^FIN")) %>% 
  group_by(course_name) %>% 
ggplot(., aes(x = score))+
  geom_histogram()+
  facet_wrap(~ course_name)+
  labs(
    title = "Exam 3 Histograms Across Classes"
  )
```

```{r}
gb_list %>% 
  group_by(course_id) %>% 
  filter(str_detect(assignment_name, "^Exam3|^EXAM3|^Fin|^FIN")) %>% 
  ggplot(.,aes(x=course_name, y = score, group=course_name))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(
    title = "Exam 3 Scores Across Different Classes",
    x = "Course",
    y = "Exam Score"
  )
```

```{r}
gb_list %>% 
  group_by(course_name) %>% 
  filter(str_detect(assignment_name, "^Exam3|^EXAM3|^Fin|^FIN")) %>%
  ggbetweenstats(
    data = .,
    x = course_id,
    y = score,
    pairwise.comparisons = FALSE)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    labs(
    title = "Exam 3 Scores Across Different Classes",
    x = "Course",
    y = "Exam Score"
  )
```
