---
title: "Home"
output: html_document
---

This site shows some ways of interacting with Canvas Learning Management System (LMS) by means of R. 

The `R\exams` package is used to generate quiz zip files that can be imported into Canvas. This is much easier and faster than entering questions into Canvas's GUI.

The `rcanvas` library is used to connect to NMSU's instance of Canvas. The primary task is retrieving grade books. From there, the data are wrangled with `dplyr`, summarized, and graphed with `ggplot2` and `ggstatsplot`.

The original R Markdown files containing the code can be accessed at the corresponding GitHub  [repository](https://github.com/n8craig/nmsu-grades).

```{r Load-libraries, include=FALSE}
# Read CSV file
library(readr)

# Data wrangling and formatting
library(tidyverse)
library(knitr)
opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=TRUE, results='asis')

# Plotting
library(ggstatsplot)
```


```{r across-course-grades, echo=FALSE, message=FALSE, warning=FALSE}

gb_list <- read_csv("data/gb_list.csv")

gb_list <- gb_list %>% 
  mutate(course_id = as_factor(course_id)) %>% 
  mutate(course_name = as_factor(course_name))

# order the factors properly
gb_list$course_name <- factor(gb_list$course_name,
                              levels= c(
                                "ANTH-125G S19",
                                "ANTH-125G F19",
                                "HON-235G F19",
                                "ANTH-125G S20",
                                "ANTH-125G F20",
                                "HON-235G F20",
                                "ANTH-1140G S21",
                                "HNRS-2161G S21"
                              ))

gb_list %>% 
  group_by(course_name) %>% 
  filter(str_detect(assignment_name, "^Exam1|^EXAM1")) %>%
  ggbetweenstats(
    data = .,
    x = course_name,
    y = score,
    pairwise.comparisons = FALSE)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    labs(
    title = "Exam 1 Scores Across Intro Anthro Classes S19-S21",
    x = "Course",
    y = "Exam Score"
  )
```

